Constructing Statistical Deformable Models with Shape Flow

### 2015-07-16 contents/implementation.tex
changes tassos shapeflow, locally

%two major modules. The first module handles inconsistent annotation set by converting to landmark independent shape discriminator. While the other module produces shape flow on object discriminators to generate dense flow transformations in shape space following by robust PCA\cite{?} to generate deformable model. In this section, we present the entire architectures, design decision and algorithms.

\subsection{Shape Representation Without Consistent Annotations}
\label{sec:svs}
%Ordinarily, construction of deformable model requires training data set having consistent number of landmarks. But unavoidable changes has to be made before applying same algorithm on diverse annotated data set.

@@

Using the SVS functions built from training data, the next step of our pipeline is to establish dense correspondences between all these functions. We propose to do this procedure jointly, by using the correspondence basis. This procedure shares similarities with multiframe optical flow and we call it Shape flow. We build upon robust methods for  optical flow estimation. Optical flow estimation typically works based on the assumptions of brightness or colour constancy and motion smoothness. However, in terms of shapes, neither constant illumination nor continuous motion exists in sequence of shapes as all training data are independent. 
For this reason, we propose to modify the formulation of multiframe optical flow by using the correspondence basis that we introduced in conjunction with the SVS functions, rather than actual images.

\subsubsection{Non-rigid ICP and Correspondence Basis Creation} \label{sec:trabasis}
Since shapes from training data set having completely no motion smoothness, it is of importance to contain constrains in optimization on how pixels moves from one shape frame to another. We constrain the optimization on correspondence subspace.
We supervise trajectories from sparse landmarks. To begin with, Non-rigid Iterative Closest Point (NICP)\cite{Amber2007} applied on sparse annotated points. NICP deforms target points to fit the template iteratively until all points on target deformed to match points on template. Performs the algorithm for all training data will get an implicit point correlation between sparse shapes.

Since optical flow pixel-wise frame registration, the trajectory basis build should also match the dimension, which is dense on shapes with trajectory length depends on number of training data. With correspondences produced after applying NICP, Thin-Plate Spline (TPS)\cite{Bookstein1989} transformations are utilized to warp every shape planes to template plane thereby generates an implicit dense shape registering. For all dense transformations $\bm{u_n}(\bm{x}), n \in \{1,...,F\}$, where $F$ is number of data and $\bm{x}$ is vector of pixels, Principle Component Analysis (PCA) is performed on trajectory to obtain low rank trajectory basis:
\begin{equation}
    \begin{bmatrix}
        \bm{u_1}(\bm{x}) \\
        \vdots \\
        \bm{u_F}(\bm{x})
    \end{bmatrix}
    =
    \begin{bmatrix}
        \bm{q_1}(1) & \cdots & \bm{q_R}(1) \\
        \vdots      & \ddots & \vdots  \\
        \bm{q_1}(F) & \cdots & \bm{q_R}(F)
    \end{bmatrix}
    \times
    \begin{bmatrix}
        \bm{v_1}(x) \\
        \vdots \\
        \bm{v_R}(x)
    \end{bmatrix}
\end{equation}
where $\bm{q_i}(n)$ are low rank components with $R \ll 2F$ and $\bm{v_i}(x)$ weighted each component with dependencies on $x$. Simpler expression shown below:
\begin{equation}
    \bm{u_n}(\bm{x})=\sum_{i=1}^R\bm{q_i}(n)\bm{v_i}(x)+\bm{\varepsilon_n}(\bm{x})
\end{equation}
Although optical flow on shapes made under the assumption that objects motion are smooth, the low rank constrain on trajectory basis recovered the hypothesis.

>>>>

%two major modules. The first module handles inconsistent annotation set by converting to landmark independent shape discriminator. While the other module produces shape flow on object discriminators to generate dense flow transformations in shape space following by robust PCA\cite{?} to generate deformable model. In this section, we present the entire architectures, design decision and algorithms.
\subsection{Shape Representation Without Consistent Annotations}
\label{sec:svs}
%Ordinarily, construction of deformable model requires training data set having consistent number of landmarks. But unavoidable changes has to be made before applying same algorithm on diverse annotated data set.

@@

Using the SVS functions built from training data, the next step of our pipeline is to establish dense correspondences between all these functions. We propose to do this procedure jointly, by using the correspondence basis. This procedure shares similarities with multiframe optical flow estimation \cite{Garg:2013hu} and we call it Shape flow. We build upon robust methods for  optical flow estimation. Optical flow estimation typically works based on the assumptions of brightness or colour constancy and motion smoothness. However, in terms of shapes, neither constant illumination nor continuous motion exists in sequence of shapes as all training data are independent.
For this reason, we propose to modify the formulation of multiframe optical flow by using the correspondence basis that we introduced in conjunction with the SVS functions, rather than actual images.

\subsubsection{Non-rigid ICP and Correspondence Basis Creation} \label{sec:trabasis}

Instead of the \emph{motion} basis used in multiframe optical flow formulation of \cite{Garg:2013hu}, we build a \emph{correspondence} basis that introduces constrains in how points of different shapes are matched to each other. By considering a reference shape, every pixel of the image that represents this shape is matched with its corresponding position and in this way defines a \emph{correspondence vector}. This vector consists of the 2D locations of the specific point in all shape images. To form this vector, the training shapes are arranged in an arbitrary order. Similarly to the order of training samples when Principal Component Analysis (PCA) is applied, this order does not affect the result of our method and any re-ordering would produce exactly the same results.

The correspondence basis that we build constrains these correspondence vectors to lie on a subspace and in this way acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes. To effectively build this basis, we first apply Non-rigid Iterative Closest Point (NICP)\cite{Amber2007} between the cloud of sparse annotated landmarks in the reference shape and the one of every shape of the training set. NICP deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. To make this correspondence vectors estimation dense (i.e.~for every pixel of the reference shape), we apply Thin-Plate Splines (TPS)\cite{Bookstein1989} interpolation. This initial estimation of dense correspondence vectors define for every training shape a warping function that registers it with the reference shape and equivalently an initial ``shape flow'', i.e.~ an ``optical flow'' between the reference shape and every training shape.

The correspondence basis is found by applying PCA on the result of TPS interpolation and keeping only the first $R$ principal components. Formally, let $F$ be the number of training shapes and $n=1,\ldots,F$ be the training shape index. Also, let $\boldsymbol{q}_1(n),\ldots,\boldsymbol{q}_R(n):
\{1,\ldots,F\} \rightarrow \R^2$ be the $R$ orthonormal elements of the correspondence basis, where $\boldsymbol{q}_i(n)$ is the displacement vector that matches the points of the reference shape with the points of the $n$-th training shape shape, according to the variation described from the $i$-th correspondence basis element. Note that the basis elements $\boldsymbol{q}_i(n)$ are independent from the point location.

In addition, let .. be the image domain where the reference and training shapes are defined and ... denote the point location. We denote the shape flow result as ...,  where u_n(x) is the displacement vector that matches the point ... of the image of the reference shape with its corresponding location at the image of the $n$-th training shape.

Using the constructed correspondence basis, the shape flow can be approximated as:

....

where $\bm{v_i}(\bm{x})$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as linear combination of basis elements that is controlled by the coefficients $\bm{v_i}(\bm{x})$ that depend on $\bm{x}$.
The values of the $i$-th coefficient for all the points $\bm{v_i}(\bm{x})$ can be interpreted as an image defined on \Omega. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $\bm{v_i}(\bm{x})$

, where

We apply PCA on the

a set of warping functions that give an initial registration of every training shape with the reference shape.

Performs the algorithm for all training data will get an implicit point correlation between sparse shapes.

Since optical flow pixel-wise frame registration, the trajectory basis build should also match the dimension, which is dense on shapes with trajectory length depends on number of training data. With correspondences produced after applying NICP, Thin-Plate Spline (TPS)\cite{Bookstein1989} transformations are utilized to warp every shape planes to template plane thereby generates an implicit dense shape registering. For all dense transformations $\bm{u_n}(\bm{x}), n \in \{1,...,F\}$, where $F$ is number of data and $\bm{x}$ is vector of pixels, Principle Component Analysis (PCA) is performed on trajectory to obtain low rank trajectory basis:
\begin{equation}
    \begin{bmatrix}
        \bm{u_1}(\bm{x}) \\
        \vdots \\
        \bm{u_F}(\bm{x})
    \end{bmatrix}
    =
    \begin{bmatrix}
        \bm{q_1}(1) & \cdots & \bm{q_R}(1) \\
        \vdots      & \ddots & \vdots  \\
        \bm{q_1}(F) & \cdots & \bm{q_R}(F)
    \end{bmatrix}
    \times
    \begin{bmatrix}
        \bm{v_1}(x) \\
        \vdots \\
        \bm{v_R}(x)
    \end{bmatrix}
\end{equation}
where $\bm{q_i}(n)$ are low rank components with $R \ll 2F$ and $\bm{v_i}(x)$ weighted each component with dependencies on $x$. Simpler expression shown below:
\begin{equation}
    \bm{u_n}(\bm{x})=\sum_{i=1}^R\bm{q_i}(n)\bm{v_i}(\bm{x})+\bm{\varepsilon_n}(\bm{x})
\end{equation}
Although optical flow on shapes made under the assumption that objects motion are smooth, the low rank constrain on trajectory basis recovered the hypothesis.
